Generating urban morphologies at large scales

### 2019-03-15 paper.tex
.

We compute the morphological indicators given above on a large sample on real urban areas. For practical computational reasons, we restrain our geographical area of study to European urban areas as provided by \cite{}. We expect to already have a good representativity, although not universal, of existing urban forms with this sample, as it is known that European cities already have a significant morphological diversity \citep{le2015forme}. We collect building layouts from OpenStreetMap (illustrated in Figure~\ref{fig:osm}), as this source has been shown to have a good quality especially in Europe \citep{mooney2010towards}. Using the \texttt{osmosis} tool, buildings are filtered from the openstreetmap raw dump for Europe (downloaded from \url{http://download.geofabrik.de/} in March 2019) and inserted into a Postgis database, which can then be efficiently queried for a specific bounding box. Indeed, although the library developed provides a direct access to the OpenStreetMap API, query limitations do not allow such a systematic sampling.
%Our tool also provide a direct access to the classic and overpass API % not necessary ?
We sample $N=72,000$ points into polygons corresponding to urban areas, first by selecting the area with a uniform selection weighted by population of areas, then by drawing uniform spatial coordinates within the polygon with a polygon sampling heuristic. % TODO clarify the polygon sampler

@@

%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.767548582324828,42.12774259047379)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.956412030079216,48.95833438674378)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.8534915107882908,48.87943398394085)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(4.724576799597424,46.11247464316661)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(6.693161940127149,51.28067900621835)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(7.665417780766332,51.61719462734998)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(7.746400123854242,49.448716515580514)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(8.66720808445827,49.469125872820456)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(8.368921085687871,49.224787816959754)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(9.3988783289936,45.52643666558245)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(9.422039749719488,48.68506062982569)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(10.529303821177011,57.458935658684915)}.png}
    \end{subfigure}
    \caption{Samples extracted from OpenStreetMap.}
    % TODO should provide parameters here
    \label{fig:osm}
\end{figure}
%%%%%%%%%%%%%%

@@

To obtain typical representative points, we proceed to an unsupervised clustering on the two first principal components of these points. Using a k-means algorithm (5000 stochastic repetitions), varying the number of clusters shows an endogenous transition in the within-cluster variance proportion, suggesting to take $k=4$.
Examples within the classes are shown in Fig.~\ref{fig:osm_types}.
The centroids will be used as typical objectives for model calibration.

We can also consider the distribution of these measures within sampled urban areas.

%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \rotatebox{90}{Class 1}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_2.43416877400257_48.5494220403648_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_3.20479601880545_50.1572174642354_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_6.03454192624923_43.1438923676911_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_7.76412991205779_47.6437664789465_1}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 2}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_4.73115258198923_45.9763617017002_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_4.92601740592234E-5_50.7945845258906_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_8.76440233864255_50.3459471560717_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_10.6351277289413_51.1689134935896_2}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 3}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_2.78910884212325_50.5481512216399_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_20.0294237103859_51.5168560281418_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_5.55350495469112_51.4926898192205_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_23.8025598321731_54.950616933903_3}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 4}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_5.28978808584814_52.0605158929046_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_6.11373040780706_50.9165373258647_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_8.35334588815557_49.530653800312_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_12.4396942996003_50.983284784414_4}.png}
  \end{subfigure}\\
  \caption{Typology applied to OpenStreetMap samples.}
  % TODO should provide parameters here
  \label{fig:osm_types}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%
\subsection{Direct sampling}

@@

%%%%%%%%%%%%%%
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{lhscalib_projrealpcs}
    % TODO figure a refaire + corresponding examples of morphologies
    \caption{Comparison of real morphologies with patterns produced by the synthetic generators. All points are projected into the first two components computed on the real measures only. Points in dark blue correspond to real configurations. The type of generator is given by the color in legend. Dark circle correspond to the centroids of clusters on which the calibration is done.}
    \label{fig:lhs}
\end{figure*}
%%%%%%%%%%%%%%

>>>>

We compute the morphological indicators given above on a large sample on real urban areas. For practical computational reasons, we restrain our geographical area of study to European functional urban areas as provided by \cite{bretagnolle2019following}. We expect to already have a good representativity, although not universal, of existing urban forms with this sample, as it is known that European cities already have a significant morphological diversity \citep{le2015forme}. We collect building layouts from OpenStreetMap (illustrated in Figure~\ref{fig:osm}), as this source has been shown to have a good quality especially in Europe \citep{mooney2010towards}. Using the \texttt{osmosis} tool, buildings are filtered from the openstreetmap raw dump for Europe (downloaded from \url{http://download.geofabrik.de/} in March 2019) and inserted into a Postgis database, which can then be efficiently queried for a specific bounding box. Indeed, although the library developed provides a direct access to the OpenStreetMap API, query limitations do not allow such a systematic sampling.
%Our tool also provide a direct access to the classic and overpass API % not necessary ?
We sample $N=72,000$ points into polygons corresponding to urban areas, first by selecting the area with a uniform selection weighted by population of areas, then by drawing uniform spatial coordinates within the polygon with a polygon sampling heuristic. % TODO clarify the polygon sampler

@@

%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.767548582324828,42.12774259047379)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.956412030079216,48.95833438674378)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(2.8534915107882908,48.87943398394085)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(4.724576799597424,46.11247464316661)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(6.693161940127149,51.28067900621835)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(7.665417780766332,51.61719462734998)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(7.746400123854242,49.448716515580514)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(8.66720808445827,49.469125872820456)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(8.368921085687871,49.224787816959754)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(9.3988783289936,45.52643666558245)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(9.422039749719488,48.68506062982569)}.png}
    \end{subfigure}
    \begin{subfigure}[b]{0.1\textwidth}
        \includegraphics[width=\textwidth]{{osm_(10.529303821177011,57.458935658684915)}.png}
    \end{subfigure}
    \caption{Samples extracted from OpenStreetMap.}
    % provide coordinates ?
    \label{fig:osm}
\end{figure}
%%%%%%%%%%%%%%

@@

To obtain typical representative points, we proceed to an unsupervised clustering on the two first principal components of these points. Using a k-means algorithm (5000 stochastic repetitions), varying the number of clusters shows an endogenous transition in the within-cluster variance proportion, suggesting to take $k=4$.
Examples within the classes are shown in Fig.~\ref{fig:osm_types}. %TODO discuss a bit
The centroids will be used as typical objectives for model calibration.

We can also consider the distribution of these measures within sampled urban areas. Keeping the areas with more than 10 sample points, we obtain 219 areas, for which we can compute the proportion of points within each morphological cluster. An Herfindhal diversity index on these proportions $p_k \in \left[0,1\right]$ computed as $h = 1 - \sum_k p_k^2$ ranges between 0.31 and 0.75 with an average of 0.63, suggesting very different profiles of urban areas. A chi-squared test of 10 levels of this diversity with the country is not significant (p=0.7), but the diversity index negatively correlates ($\rho = -0.11$, Fisher 95\% confidence interval $\left[-0.24,0.02\right]$) with longitude, meaning that Western cities are more diverse than Eastern cities, and more slightly with latitude ($\rho = 0.08 \left[-0.04,0.21\right]$).

%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \rotatebox{90}{Class 1}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_2.43416877400257_48.5494220403648_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_3.20479601880545_50.1572174642354_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_6.03454192624923_43.1438923676911_1}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/1/osm_7.76412991205779_47.6437664789465_1}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 2}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_4.73115258198923_45.9763617017002_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_4.92601740592234E-5_50.7945845258906_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_8.76440233864255_50.3459471560717_2}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/2/osm_10.6351277289413_51.1689134935896_2}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 3}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_2.78910884212325_50.5481512216399_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_20.0294237103859_51.5168560281418_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_5.55350495469112_51.4926898192205_3}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/3/osm_23.8025598321731_54.950616933903_3}.png}
  \end{subfigure}\\
  \rotatebox{90}{Class 4}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_5.28978808584814_52.0605158929046_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_6.11373040780706_50.9165373258647_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_8.35334588815557_49.530653800312_4}.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.1\textwidth}
    \includegraphics[width=\textwidth]{{osm/4/osm_12.4396942996003_50.983284784414_4}.png}
  \end{subfigure}\\
  \caption{Typology applied to OpenStreetMap samples.}
  % provide coordinates ?
  \label{fig:osm_types}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%
\subsection{Model simulation and calibration}

@@

%%%%%%%%%%%%%%
\begin{figure*}
\vspace{-1.5cm}
    \centering
    (1)
    \framebox{\includegraphics[width=0.18\linewidth]{closest1_blocks}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest1_real}}\hspace{0.4cm}
    (2)
    \framebox{\includegraphics[width=0.18\linewidth]{closest2_expMixture}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest2_real}}\\\vspace{0.2cm}
    (3)
    \framebox{\includegraphics[width=0.18\linewidth]{closest3_blocks}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest3_real}}\hspace{0.4cm}
    (4)
    \framebox{\includegraphics[width=0.18\linewidth]{closest4_percolation}}
    \framebox{\includegraphics[width=0.18\linewidth]{closest4_real}}\\
    \includegraphics[width=\linewidth]{lhscalib_projrealpcs}
    \caption{Comparison of real morphologies with patterns produced by the synthetic generators. All points are projected into the first two components computed on the real measures only. Points in dark blue correspond to real configurations. The type of generator is given by the color in legend. Dark circle correspond to the centroids of clusters on which the calibration is done, their number giving the corresponding configurations shown above. For each, we successively show the best synthetic configuration ((1) block generator, (2) exponential mixture generator, (3) block generator, (4) percolation generator), and the real configuration from OSM.}
    \label{fig:lhs}
\end{figure*}
%%%%%%%%%%%%%%
